---
title: "Recurrence Prediction Model"
format: html
---

# 1. Problem definition

This project amis to predict whether an individual will experience a recurrence event within 180 days, based on information from their first episode. The goal is to explore how different feature representations affect model performance and interpretability.

---

# 2. Data simulation

The dataset is simulated to reflect a simplified event-based system.

- Each individual is assigned a risk level (low, medium, high)
- Recurrence probability is driven primarily by risk level
- Episode duration is generated independently

```{r}

source("../R/simulate_data.R")

df <- simulate_events()

head(df)
```

# 3. Feature engineering

Two types of features are created:

- Continuous: episode duration
- Categorical: episode type (short, medium and long)

```{r}

source("../R/feature_engineering.R")

df_feat <- create_features(df)

head(df_feat)

```

# 4. Model specification

Two logistic regression models are compared:

- Model A: continuous duration
- Model B: categorical duration

```{r}

source("../R/model.R")

model_cont <- fit_model_continuous(df_feat)
model_cat <- fit_model_categorical(df_feat)

```

# 5. Model evaluation

```{r}

df_pred_cont <- add_prediction(df_feat, model_cont)
res_cont <- evaluate_model(df_pred_cont)

df_pred_cat <- add_prediction(df_feat, model_cat)
res_cat <- evaluate_model(df_pred_cat)

res_cont$accuracy
res_cat$accuracy

res_cont$baseline_accuracy

```

## 1. Recurrence rate by risk level

Recurrence increases clearly with risk level, confirming that risk is the primary driver in this dataset.

```{r}

library(ggplot2)

df_plot <- df_feat |>
  group_by(risk_level) |>
  summarise(recur_rate = mean(y_recur_within_180),
            n = n())

ggplot(df_plot,
       aes(x = risk_level,
           y = recur_rate)
       ) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Recurrence rate by risk level",
       x = "Risk level",
       y = "Recurrence %")

```

## 2. Predicted probability distribution

The model assigns higher predicted probabilities to actual recurrence cases, indicating it captures meaningful signal.

```{r}

ggplot(df_pred_cont,
       aes(x = pred_prob,
           fill = factor(y_recur_within_180))
       ) +
  geom_histogram(position = "identity",
                 alpha = 0.5,
                 bins = 30) +
  labs(title = "Predicted probability distribution",
       x = "Predicted probability",
       fill = "Actual outcome")

```

# 6. Key findings

Both models produced similar performance. This suggests that:

- Risk level is the primary driver of recurrence
- Episode duration (both continuous and categorical) does not add meaningful predictive signal in this setup

# 7. Interpretation

This result is expected given the data generating process:

- Recurrence probability is determined by risk level
- Episode duration is randomly generated

Therefore, duration acts as noise rather than signal.

# 8. Conclusion

This project demonstrates:

- How to structure a prediction problem from event data
- The imporance of feature design and representation
- The need to evaluate models against baseline performance

Even when additional features are included, they must carry real signal to improve predictive performance.

This is a modelling project that explores how feature representation affects recurrence prediction, highlighting the importance of signal versus noise.


